SoK: Application Attack Defenses (in Question)

### 2014-09-23 docs/questioning.tex
small caps where needed, fixed two typos

As in medical diagnostic tests, application attack defense mechanisms
must demonstrate the presence of an attack. This, however, does not
make an application attack defense mechanism immediately useful. In
order for a detection mechanism to be useful it must accurate, easy to
use, and economical. The accuracy of a detection mechanism is gauged
with the following metrics:
\begin{itemize}
\item {\bf Sensitivity}, the probability that an attack will be
  caught.
\item {\bf Specificity}, the probability that a normal interaction
  will test negative.
\item {\bf Positive Predictive Value} (PPV), the probability that a
  reported attack is a real attack.
\item {\bf Negative Predictive Value} (NPV), the probability that if
  nothing is reported no attack has taken place.
\end{itemize}

For example, a simple announcement that an attach defense mechanism
catches an attack says very little about the usability and
applicability of the mechanism. There may be some value in indicating
that detection \emph{per se} is possible, if that has been considered
infeasible in the past. Most value, though, comes from knowing the
answers to questions like the following, which correspond to the
previous metrics:
\begin{itemize}
\item Does it catch all the attacks? If not, how many of them?
\item Does it catch only attacks? If not, how many false alarms would
  we expect?
\item How much should we be worried if we do get an alarm?
\item How much at ease should we be if no alarm is raised?
\end{itemize}

Sensitivity, specificity, PPV, and NPV, are defined precisely by way
of the $2\times 2$ Table~\ref{tab:attacks-outcomes}. In the cells of
the table we distinguish:
\begin{itemize}
\item True Positive (TP), a real attack that raises an alarm.
\item False Positive (FP), an event that although it is not an attack
  raises an alarm.
\item True Negative (TN), an event that is not an attack and that does
  not raise an alarm.
\item False Negative (FN), an event that although is as attack does
  not raise an alarm.
\end{itemize}

With these we can calculate:

\begin{equation}
\textrm{sensitivity} = \frac{\textrm{TP}}{\textrm{TP} + \textrm{TN}}
\end{equation}

\begin{equation}
\textrm{specificity} = \frac{\textrm{TP}}{\textrm{TN} + \textrm{FP}}
\end{equation}

\begin{equation}
\textrm{PPV} = \frac{\textrm{TP}}{\textrm{TP} + \textrm{FP}}
\end{equation}

\begin{equation}
\textrm{NPV} = \frac{\textrm{TN}}{\textrm{TN} + \textrm{FN}}
\end{equation}

@@

The calculation of PPV and NPV requires that we evaluate the attack
defense mechanism on a real operational environment where the events
are a mix of attacks and innocuous events and, moreover, the mix ratio
is the ratio of attacks over harmless events that exists in actual
operations. Too few or too many attacks will skew metrics calculations.

The last requirement may seem cumbersome, but consider that, in actual
operations, PPV is what would determine our actions. Administrators
will limited resources need to allocate their attention judiciously.
They may need to prioritize responses to emergency warnings; the PPV
value will indicate them what priority to give to a raised alarm.

>>>>

As in medical diagnostic tests, application attack defense mechanisms
must demonstrate the presence of an attack. This, however, does not
make an application attack defense mechanism immediately useful. In
order for a detection mechanism to be useful it must be accurate, easy to
use, and economical. The accuracy of a detection mechanism is gauged
with the following metrics:
\begin{itemize}
\item {\bf Sensitivity}, the probability that an attack will be
  caught.
\item {\bf Specificity}, the probability that a normal interaction
  will test negative.
\item {\bf Positive Predictive Value} ({\sc ppv}), the probability that a
  reported attack is a real attack.
\item {\bf Negative Predictive Value} ({\sc npv}), the probability that if
  nothing is reported no attack has taken place.
\end{itemize}

For example, a simple announcement that an attack defense mechanism
catches an attack says very little about the usability and
applicability of the mechanism. There may be some value in indicating
that detection \emph{per se} is possible, if that has been considered
infeasible in the past. Most value, though, comes from knowing the
answers to questions like the following, which correspond to the
previous metrics:
\begin{itemize}
\item Does it catch all the attacks? If not, how many of them?
\item Does it catch only attacks? If not, how many false alarms would
  we expect?
\item How much should we be worried if we do get an alarm?
\item How much at ease should we be if no alarm is raised?
\end{itemize}

Sensitivity, specificity, {\sc ppv}, and {\sc npv}, are defined precisely by way
of the $2\times 2$ Table~\ref{tab:attacks-outcomes}. In the cells of
the table we distinguish:
\begin{itemize}
\item True Positive ({\sc tp}), a real attack that raises an alarm.
\item True Negative ({\sc tn}), an event that is not an attack and that does
  not raise an alarm.
\item False Positive ({\sc fp}), an event that although it is not an attack
  raises an alarm.
\item False Negative ({\sc fn}), an event that although is as attack does
  not raise an alarm.
\end{itemize}

\noindent
With these we can calculate:

\begin{equation}
\textrm{Sensitivity} = \frac{\textrm{{\sc tp}}}{\textrm{{\sc tp}} + \textrm{{\sc tn}}}
\end{equation}

\begin{equation}
\textrm{Specificity} = \frac{\textrm{{\sc tp}}}{\textrm{{\sc tn}} + \textrm{{\sc fp}}}
\end{equation}

\begin{equation}
\textrm{{\sc ppv}} = \frac{\textrm{{\sc tp}}}{\textrm{{\sc tp}} + \textrm{{\sc fp}}}
\end{equation}

\begin{equation}
\textrm{{\sc npv}} = \frac{\textrm{{\sc tn}}}{\textrm{{\sc tn}} + \textrm{{\sc fn}}}
\end{equation}

@@

The calculation of {\sc ppv} and {\sc npv} requires that we evaluate the attack
defense mechanism on a real operational environment where the events
are a mix of attacks and innocuous events and, moreover, the mix ratio
is the ratio of attacks over harmless events that exists in actual
operations. Too few or too many attacks will skew metrics calculations.

The last requirement may seem cumbersome, but consider that, in actual
operations, {\sc ppv} is what would determine our actions. Administrators
will limited resources need to allocate their attention judiciously.
They may need to prioritize responses to emergency warnings; the {\sc ppv}
value will indicate them what priority to give to a raised alarm.
